Moving Together

### 2023-03-14 graphs_newt_snakes/move_together.Rmd
got some writting done for this markdown.

What does it mean to be correlated? Might this occur if I simulate some controlled movement?

@@

Each GA combination and trial has its own msprime simulation, but the msprime file is shared between the 5 section and 7 section slim runs.

### The data

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

>>>>

What does it mean to be correlated? Might this occur if I simulate some semi-controlled movement?

@@

Each GA combination and trial has its own msprime simulation. I start both the newt and snake population in the bottom third section of my map. Over time they eventually disperse into the entire map. The background enverionment is homogeneous.

@@

There are some interesting differences taking place here, when compared to simulation without movement and maps. It seems like it takes a bit longer to reach a steady state (but this could also be the stochasticity of the simulation). The mean newt and snake phenotype get to similar values as my original simulations. Still follows what I have seen before, the mean phenotypes of newts and snakes go up, until they reach an equilibrium.

@@

These results are similar to what I have seen before. There might be an increase in the differences between snake and newt phenotype, but the overall message higher phenotype = more individuals in unchanged. However, the layout of the points seems to be flatter and stretched.

@@

The spatial phenotype correlation seems very random.

@@

It seems like there is an increase in phenotype spatial correlation near the beginning of the simulation (<25,000 generations), but as time goes on it is lost. The spatial phenotype correlation is also lower than the spatial correlation that result from simulations run with a gradient map.

@@

Most of these plots show that as time increases both the mean newt & snake phenotype. Some of these plots have an increase in the spatial phenotype correlation (pink line), but this increase in the correlation between local newt and snake phenotypes can also seem very random. Sometimes the spatial phenotype correlation increases and then decreases overtime. Under certain GA combinations (both 1e09_0.05) spatial phenotype correlation increases the most when the newt and snake mean phenotypes are increasing the most.

@@

The results looked similar to results that I have seen in other experiments, with an exception at the beginning. In the beginning of the simulation both newt and snake population grows, but this growth has a longer tail do to the populations expanding territory. The population size reaches a steady point and then newts and snakes co-evolve. In the middle part of my simulation, the difference between newt and snakes mean phenotype solidifies. Often, the difference in mean phenotype decreases (where compared to the beginning of the simulation). When the GA has a high mutation rate and low mutation effect size (GA 1), the difference in mean phenotype grows. This leads to the species with GA 1 losing the co-evolutionary arms race. More frequent smaller steps does not help a species win in an arms race (might also be due to lower mutational variance). The histograms reflect what is seen in the scatter plots.

@@

Overall, these plots are very similar to the plots seen with and without a gradient map. The largest difference is seen from plots at the beginning of the simulation. At the beginning of the simulation the points are more spread out (less in a line), eventually they form the line of shapes and colors seen in my previous simulations. There are clusters of points that create lines of shapes and lines of colors. The best GA for newts and snakes was 1e-09_0.05 (can be seen as green near the top of all the lines of shapes and triangles near the bottom right look at all the lines of color). It is interesting to see how these points spread apart, but remain similar between trials. When putting these figures together it seems like the population size of snakes is lower when the newt phenotype is larger than the snake phenotype.

@@

In the heatmap plots each GA combination and trails is presented by combining newt GA in the x-axis to snake GA and trial number in the y-axis. The result is the color in that section. There are two types of heatmap plots shown below. One shows the average snake population size for a time chunk with darker colors indicating a smaller snake population and lighter colors indicating a larger snake population. The other heatmap shows the average difference between snake and newt mean phenotype (red=newts had a higher phenotype, blue=snakes had a higher phenotype). I look at 3 time slices for both types of heatmaps

@@

As with results seen in the summary plots, the heat map plots tell a similar story. At the beginning of the simulation the snake population size and the mean difference between snake and newt phenotype varies a lot between the simulation trials, more than any other simulation that I have currently run. But by the end of the simulation the results remain similar to all of the other simulations that I have run.  Theses plots shows that under certain GA either newts or snakes have an co-evolutionary advantage (best GAs 1e-9(0.05)^2 = 2.5e-12 or 1e-10(0.5)^2 = 2.5e-11).

@@

These results are interesting, they show that movement might create an increase in spatial phenotype correlation, but that this correlation disappears over time. Very interesting results.
